---
title: "Movies"
author: "Var Kurapati"
execute:
  echo: false
---

Rotten Tomatoes publishes two scores for every movie: the **Tomatometer**, which 
is the percentage of approved critics who gave a positive review, and the 
**audience score**, which comes from verified ticket buyers who rate films on a 
0–5 scale converted to a percentage. The same movie can score very differently 
on each scale. This analysis uses a dataset of 1,000 movies to ask: do critics 
and audiences systematically disagree, and which genres drive the biggest gap 
between their scores?

```{r}
#| message: false
library(tidyverse)
```

```{r}
#| message: false
movies <- read_csv("data/rotten_tomatoes_movies.csv", show_col_types = FALSE)
```

```{r}
#| cache: true
movies_long <- movies |>
  separate_longer_delim(genre, delim = ", ") |>
  filter(!is.na(genre), genre != "")
```

```{r}
movies_gap <- movies |>
  mutate(gap = tomatoMeter - audienceScore) |>
  filter(!is.na(gap))
```

```{r}
gap_by_genre <- movies_gap |>
  separate_longer_delim(genre, delim = ", ") |>
  filter(!is.na(genre), genre != "") |>
  group_by(genre) |>
  summarise(
    mean_gap = mean(gap, na.rm = TRUE),
    n = n()
  ) |>
  filter(n >= 50) |>
  arrange(mean_gap)
```

```{r}
gap_by_genre |>
  ggplot(aes(x = mean_gap, y = fct_reorder(genre, mean_gap), fill = mean_gap > 0)) +
  geom_col() +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  scale_fill_manual(values = c("TRUE" = "#2196F3", "FALSE" = "#F44336"),
                    labels = c("TRUE" = "Critics rate higher", "FALSE" = "Audiences rate higher")) +
  labs(
    title = "Action and family films win audiences; documentaries win critics",
    subtitle = "Mean difference in Tomatometer minus audience score, genres with 50+ movies",
    x = "Mean gap (critic score minus audience score)",
    y = NULL,
    fill = NULL,
    caption = "Source: Rotten Tomatoes via Kaggle (stefanoleone992)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```





